An example of extremely reproducible enrichment analysis

### 2023-06-22 example.Rmd
minor adjustment to euler diagram
bunch of changes

@@

Get the gene sets loaded in R.

```{r,gs}
# from https://reactome.org/download/current/ReactomePathways.gmt.zip
genesets2 <- read.gmt("ReactomePathways_2023-03-06.gmt")
gsets <- gmtPathways("ReactomePathways_2023-03-06.gmt")

@@

@@

Enrichment firstly with upregulated genes

@@

ora_up <- as.data.frame(enricher(gene = defup ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

@@

head(ora_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

@@

n_pw=10

ora_dn <- as.data.frame(enricher(gene = defdn ,
  universe = bg,  maxGSSize = 5000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

@@

```{r,venn1,fig.width=6}

>>>>

The gene counts have also been deposited to the `/data` folder in the `example.Rdata` file in case
the DEE2 resource becomes unavailable.
To import it, use the command: `load("data/example.Rdata")`

@@

Get the gene sets loaded in R.
They are located in the `/ref` folder of the repo.

```{r,gs}
# from https://reactome.org/download/current/ReactomePathways.gmt.zip
genesets2 <- read.gmt("ref/ReactomePathways_2023-03-06.gmt")
gsets <- gmtPathways("ref/ReactomePathways_2023-03-06.gmt")

@@

message("number of genes in each group")
lapply(list("background"=bg,"up-regulated"=defup,"down-regulated"=defdn),length)

@@

One of the quirky things I learned about clusterprofiler is that
not all genes specified by the user as background are included in
the enrichmentcalculation.
It appears as though the genes not included in the annotation set are
discarded.
I believe this is a subtle but significant problem.
To fix it is relatively easy from the user's point of view.
Just create a new gene set that consists of all detected genes and

Adding all detected genes to the background appears to improve results!

```{r,gs2}

bgdf <- data.frame("background",bg)
colnames(bgdf) <- c("term","gene")
genesets2 <- rbind(genesets2,bgdf)

Enrichment firstly with upregulated genes.
Note that the BgRatio contains the correct number of genes in the background now.

@@

# show 10 pathways only
n_pw=10

ora_up <- as.data.frame(enricher(gene = defup ,
  universe = bg,  maxGSSize = 50000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

@@

head(ora_up,n_pw) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

@@

ora_dn <- as.data.frame(enricher(gene = defdn ,
  universe = bg,  maxGSSize = 50000, TERM2GENE = genesets2,
  pAdjustMethod="fdr",  pvalueCutoff = 1, qvalueCutoff = 1  ))

@@

```{r,venn1,fig.width=8,fig.height=5}
